---
title: "Codebook Assignment"
output:
  pdf_document: default
  html_document: default
date: "November 14, 2017"
---

This assignment is meant to direct the focus of your poster project while also exposing you to the importance of a well-documented codebook. To that end you should select only the variables you are planning to include in your potential analysis, be that a linear regression model or a data visualization. 

Your final codebook document should be clearly labelled with the header titles mentioned in the following questions. For parts 2 and 3 you should write in complete sentences as if you were writing an essay. **Do not make lists.**  There should not be any code in your knitted PDF, just the output of the `codebook` function. To turn in your final assignment commit your PDF and Rmd files to Git and push them to GitHub.  

1. Create a new R Markdown file and name it after your data, followed by the word codebook (e.g.. my in-class example might be *PSID_CDI_Codebook*).

2. In the first section of your codebook, titled **Study Design**, describe the study design including the purpose of the study, the sponsor of the study, the name of the data collection organization, and the specific methodology used including the mode of data collection, the method of participant recruitment (if any), and the length of the field period. (12 points)

# Study Design

The data which I am working on, comes from different surveys. I of the data comes from Statistics Center of Iran. The data for 2006 and 2011 years are census data and almost all population is included in the census. The sponsor for all the survey and census is the Iranian Ministry of the interior. The name of data collection organization is SCI(Statistics Center of Iran). 
Data from 2001, 2008, 2009 also conducted by the same organization and same sponsor, the implementation method for these years is called sample household by the SCI. But the organization does give detailed information about the used sampling methods. The title of survey for these years is "Households Socio-economic Characteristics Survey". The surveys for 2010, 2012, 2013, and 2014 years are also sponsored by the Iranian Ministry of Interior and conducted by the SCI.  The used method for these years is survey and the purpose of the surveys is collecting data for "General Census of Population & Housing". Unfortunately,    



3. The second section, **Sampling** should clearly document all available sampling information. This includes a description of the population, the methods used to draw the sample, and any special conditions associated with the sample (i.e groups that were oversampled). (12 points)


# Sampling
The is no available information about sampling. The sections of SCI's website was unavailable to get the information. The website in Persian says that website will be unpdated,  https://www.amar.org.ir/Install/UnderConstruction.htm?aspxerrorpath=/Default.aspx 




4. Section three should be titled **Variable Index**. Here you should utilize the sample code shown in lecture (and reproduced below) to build a `data.set` version of your dataset. You will need to install and load the `memisc` package. 

Each variable in your dataset should be given a `description` and a unit of `measurement`. If there are `labels` associated with the underlying numeric values of the data those should be specified as should any `missing.value` codes including `NA`. Lastly, if your variable is a survey item/interview question you must provide the `wording` as well. 

Once you've added the information above to your `data.set` object, make a call to `codebook()` to have your variable index printed out. (26 points)   

Variable Index


```{r, echo = FALSE, message = FALSE}
# Variable Index
# Setting working directory 
library(readxl)
joint_unemployment_total <- read_excel("~/Desktop/Autumn 2017/Statistics 321/unemployment rate/joint_unemployment_total.xlsx")
joint_unemployment_total
setwd("~/desktop/Autumn 2017/statistics 321/unemployment rate")


# Loading libraries
install.packages("memisc")
install.packages("tidyverse")
library(memisc)
library(tidyverse)

# Read in data
joint_unemployment_total
```


```{r}
# Create data.set object from "data" object (joint_unemployment_total)
joint_unemployment_total
data_set <- as.data.set(joint_unemployment_total)

# Look at new data.set object
data_set

```






```{r, echo=FALSE}
# Creating variable index for data_set (truncated, more efficient code)

data_set <- within(data_set,{
  # Description of the variables
  description(gender) <- "gender is defined binary in the data and refers to men and women in the Iran"
  description(year) <- "refers to the year which data is collected"
  description(value) <- "shows the percentage of unemployment for each gender in the specified year"
  
  # Wording of survey item/interview questions (if applicable)
  wording(gender) <- "The Statistics Center of Iran did not release the questionnaire of surveys. However, in Iran mostly. I could not find the questionnaires to write working for other variables"
  
  # type of measurement
  measurement(year) <- "interval"
  measurement(value) <- "ratio"
  measurement(gender) <- "nominal"

})

```

\pagebreak

```{r}
codebook(data_set)
```




